---
title: "Lab 11 All Stars"
author: "Emily Archambault"
date: "11/8/2019"
output: html_document
---

```{r setup, message = FALSE, echo = FALSE, warning = FALSE}
library(readr)
library(dplyr)
library(babynames)
library(stringr)
library(tidyverse)

names2018 <- read_csv("yob2018.txt", 
                      col_names = FALSE, col_types = cols(X2 = col_character(),  X3 = col_integer()))
names2018 <- names2018 %>% rename("name"="X1", "sex"="X2", "n"="X3")
names2018$year <- 2018
names2018 <- names2018 %>% mutate(prop = n/sum(n))
totalnames <- rbind(babynames, names2018)
```

### Analyzing babynames

```{r, message = FALSE, echo = FALSE, warning = FALSE}
babynames_2018 <- read_csv("yob2018.txt", col_names = c("name", "sex", "n"), col_types=list(col_character(), col_character(), col_double()))
babynames_2018 <- babynames_2018 %>%
  mutate(year = 2018, prop = n/sum(n)) %>%
  select(year, name, sex, n, prop)
babynames_all <- rbind(babynames, babynames_2018)
non_gender_ambiguous <- babynames_all %>%
  spread(key=sex, value=prop) %>%
  group_by(name) %>%
  summarize(F_prop = mean(F, na.rm=TRUE), M_prop = mean(M, na.rm=TRUE)) %>%
  filter(is.na(F_prop) | is.na(M_prop) | F_prop > 10*M_prop | M_prop > 10*F_prop) %>%
  select(name, F_prop, M_prop)
  median_prop <- babynames_all %>%
  filter(year==2018) %>%
  filter(rank(-prop)>200) %>%
  summarize(max = max(prop))
trending_not_trendy <- babynames_all %>%
  filter(year == 2016|year==2018) %>%
  spread(key = year, value = prop) %>%
  mutate(one=ifelse(is.na(`2016`), 0, `2016`), three=ifelse(is.na(`2018`), 0, `2018`)) %>%
  filter(one<three) %>%
  filter(three<median_prop$max) %>%
  select(name)
cannot_shorten <- babynames_all %>%
  filter(!(str_sub(name, 1, 3) %in% babynames$name) | length(name) < 4) %>%
  select(name)
criteria <- semi_join(non_gender_ambiguous, trending_not_trendy, by="name")
criteria <- semi_join(criteria, cannot_shorten, by="name")
```

### Recommended Names

## Female -> Zarina

```{r, message = FALSE, echo = FALSE, warning = FALSE}
zarina <- totalnames %>%
  filter(str_detect(name, "^Zarina$"))
ggplot(owen) +
  geom_point(mapping=aes(x=year, y=prop)) +
  labs(title="Timeseries of 'Zarinac'") +
  xlab("Year") +
  ylab("Proporiton")
```

## Male -> Owen

```{r, message = FALSE, echo = FALSE, warning = FALSE}
owen <- totalnames %>%
  filter(str_detect(name, "^Owen$"))
ggplot(owen) +
  geom_point(mapping=aes(x=year, y=prop)) +
  labs(title="Timeseries of 'Owen'") +
  xlab("Year") +
  ylab("Proporiton")
```

We chose these names because they matched the three criteria given by the domain experts. 

### Individual Name Popularities

## Eugene Miller
```{r, message=FALSE, echo=FALSE, warning=FALSE}
#Data import & tidy
new_data <- read.delim("yob2018.txt", header = FALSE, sep = ",")
names(new_data) <- c("name", "sex", "n")
new_data <- new_data %>%
  mutate(prop = `n` / sum(new_data$n)) %>%
  mutate(year = 2018)

complete_data <- full_join(new_data,babynames)
```
```{r, message=FALSE, echo=FALSE, warning=FALSE, fig.align='center'}
#Timeseries for "Eug" and "Eugene"
my_name <- complete_data %>%
  filter(str_detect(name, "Eugene") & sex == 'M') %>%
  group_by(year) %>%
  summarise(prop = sum(prop)) %>%
  mutate("Eug" = FALSE)

initials <- complete_data %>%
  filter(str_detect(name, "^Eug") & sex == 'M') %>%
  group_by(year) %>%
  summarise(prop = sum(prop)) %>%
  mutate("Eug" = TRUE)

my_data <- full_join(my_name, initials)
my_name <- my_name %>%
  select(-c("Eug"))
initials <- initials %>%
  select(-c("Eug"))

ggplot(my_data) +
  geom_jitter(mapping = aes(x = year, y = prop, color = Eug)) + 
  labs(x = "Year", y = "Proportion", color = "First Three Letters Only", title = "Popularity of Eug and Eugene since 1880")
```

I used the string "Eugene" to search for my name, as this is the only spelling of my name I am aware of. My name seems to be the most popular name that starts with letters "Eug" as the popularity of names that start with "Eug" and "Eugene" track very closely to each other. 

```{r, message=FALSE, echo=FALSE, warning=FALSE, fig.align='center'}
#Stats on Eugene
pre_me <- my_name %>%
  filter(year == 1980)
birth_year <- my_name %>%
  filter(year == 1999)
post_me <- my_name %>%
  filter(year == 2018)

pop_data <- tribble(
  ~"Year", ~"Proportion",
  pre_me$year, pre_me$prop,
  birth_year$year, birth_year$prop,
  post_me$year, post_me$prop,
)

ggplot(pop_data) +
  geom_col(mapping = aes(x = Year, y = Proportion, fill = Proportion)) + 
  geom_text(mapping = aes(label = Year, y = Proportion/2, x = Year), color = 'white')
```

My name has become less popular since I was born, and its popularity continues to decline every year. In 1980 the popularity was `r pre_me$prop` in my birth year 1999: `r birth_year$prop` and in 2018: `r post_me$prop`. 


## Emily Archambault
 
```{r, message=FALSE, echo=FALSE, warning=FALSE}
emily <- totalnames %>% 
  filter(str_detect(name, "^E[m]+[ai][l]+[iey]+$")) %>%
  filter(sex == "F")
first3 <- totalnames %>%
  filter(str_detect(name, "^Emi")) %>%
  group_by(year)%>%
  filter(sex == "F")

ggplot() + geom_col(data = first3, mapping = aes(x = year, y = prop)) +
  geom_smooth(data = emily, mapping = aes(x = year, y = prop), se = FALSE, color = "red") + 
  annotate("text", x = 1960, y = 0.0013, label = "Emily", color = "red") +
  ggtitle("Popularity of All Female 'Emi...' Names vs Popularity of 'Emily' ")

emily18 <- emily %>%
  filter(year == "2018")
pop18 <- sum(emily18$prop)

emily01 <- emily %>%
  filter(year == "2001")
pop01 <- sum(emily01$prop)

same3 <- first3 %>%
  filter(year == "2018")
pop3 <- sum(same3$prop)
```
The regular expression I used to match the various spellings of my name is "^E[m]+[ai][l]+[iey]+$".
The popularity of my name in 2018 is `r pop18`, compared to when I was born in 2001 which was `r pop01`, a difference of
0.01107516. The popularity of all (female) names starting with "Emi" in 2018 is `r pop3`, which is 0.001304714 greater than
the proportion of my name alone.


## Jingzhi Zhang

```{r, message=FALSE, echo=FALSE, warning=FALSE}
katherine <- totalnames %>%
  filter(str_detect(name, "Katherine")) %>%
  filter(sex == 'F')
kat <- totalnames %>%
  filter(str_detect(name, "^Kat")) %>%
  group_by(year) %>%
  filter(sex == 'F')
ggplot()+
  geom_col(data = kat, mapping = aes(x = year, y = prop))+
  geom_smooth(data = katherine, mapping = aes(x = year, y = prop), se = FALSE, color = "green")+
  annotate("text", x = 1950, y = 0.0045, label = "Katherine", color = "green")
  labs(x = "Year", y = "Proportion", title = "Popularity of Female Kat and Katherine")
  
katherine2018 <- katherine %>%
  filter(year == 2018)
popular2018 <- sum(katherine2018$prop)

katherine1998 <- katherine %>%
  filter(year == 1998)
popular1998 <- sum(katherine1998$prop)

kat2018 <- kat %>%
  filter(year == 2018)
pop2018 <- sum(kat2018$prop)

kat1998 <- kat %>%
  filter(year == 1998)
pop1998 <- sum(kat1998$prop)

```

I search my name by finding the string "Katherine" directly. I found that the popularity of name "Katherine" in 2018 is `r popular2018` and the popularity of it in 1998 which is the year that I was born is `r popular1998`. From the differences of number, it seems like "Katherine" used to be a popular name but less popular at present. And the popularity of "Kat" in 2018 is `r pop2018`, compared to 1998 which is `r pop1998`, this name is much less popular than it used to be.


### Name Reccommendation
```{r, message=FALSE, echo=FALSE, warning=FALSE}
#Two uptrending names (Mia, Owen)

mia <- babynames %>%
  filter(str_detect(name,"^Mia$") & sex == "F")

ggplot(mia) +
  geom_jitter(mapping = aes(x = year, y = prop)) +
  labs(title = "Timeseries for 'Mia'")

owen <- babynames %>%
  filter(str_detect(name,"^Owen$") & sex == "M")

ggplot(owen) +
  geom_jitter(mapping = aes(x = year, y = prop)) +
  labs(title = "Timeseries for 'Owen'")
```

