---
title: "Lab 11 All Stars"
author: "Johnny Tamanaha, Emily Archambault, Kaiyuan Fan, Eugene Miller, Jingzhi Zhang"
date: "11/8/2019"
output: html_document
---

```{r setup, message = FALSE, echo = FALSE, warning = FALSE}
library(readr)
library(dplyr)
library(babynames)
library(stringr)
library(tidyverse)

names2018 <- read_csv("yob2018.txt", 
                      col_names = FALSE, col_types = cols(X2 = col_character(),  X3 = col_integer()))
names2018 <- names2018 %>% rename("name"="X1", "sex"="X2", "n"="X3")
names2018$year <- 2018
names2018 <- names2018 %>% mutate(prop = n/sum(n))
totalnames <- rbind(babynames, names2018)
```

### Analyzing babynames

```{r, message = FALSE, echo = FALSE, warning = FALSE}
babynames_2018 <- read_csv("yob2018.txt", col_names = c("name", "sex", "n"), col_types=list(col_character(), col_character(), col_double()))
babynames_2018 <- babynames_2018 %>%
  mutate(year = 2018, prop = n/sum(n)) %>%
  select(year, name, sex, n, prop)
babynames_all <- rbind(babynames, babynames_2018)
non_gender_ambiguous <- babynames_all %>%
  spread(key=sex, value=prop) %>%
  group_by(name) %>%
  summarize(F_prop = mean(F, na.rm=TRUE), M_prop = mean(M, na.rm=TRUE)) %>%
  filter(is.na(F_prop) | is.na(M_prop) | F_prop > 10*M_prop | M_prop > 10*F_prop) %>%
  select(name, F_prop, M_prop)
  median_prop <- babynames_all %>%
  filter(year==2018) %>%
  filter(rank(-prop)>200) %>%
  summarize(max = max(prop))
trending_not_trendy <- babynames_all %>%
  filter(year == 2016|year==2018) %>%
  spread(key = year, value = prop) %>%
  mutate(one=ifelse(is.na(`2016`), 0, `2016`), three=ifelse(is.na(`2018`), 0, `2018`)) %>%
  filter(one<three) %>%
  filter(three<median_prop$max) %>%
  select(name)
cannot_shorten <- babynames_all %>%
  filter(!(str_sub(name, 1, 3) %in% babynames$name) | length(name) < 4) %>%
  select(name)
criteria <- semi_join(non_gender_ambiguous, trending_not_trendy, by="name")
criteria <- semi_join(criteria, cannot_shorten, by="name")
```

### Recommended Names

## Female -> Zarina

```{r, message = FALSE, echo = FALSE, warning = FALSE}
zarina <- totalnames %>%
  filter(str_detect(name, "^Zarina$"))
ggplot(zarina) +
  geom_point(mapping=aes(x=year, y=prop)) +
  labs(title="Timeseries of 'Zarina'") +
  xlab("Year") +
  ylab("Proportion")
```

## Male -> Owen

```{r, message = FALSE, echo = FALSE, warning = FALSE}
owen <- totalnames %>%
  filter(str_detect(name, "^Owen$"))
ggplot(owen) +
  geom_point(mapping=aes(x=year, y=prop)) +
  labs(title="Timeseries of 'Owen'") +
  xlab("Year") +
  ylab("Proportion")
```

We chose these names because they matched the three criteria given by the domain experts. 

### Individual Name Popularities

## Eugene Miller
```{r, message=FALSE, echo=FALSE, warning=FALSE}
#Data import & tidy
new_data <- read.delim("yob2018.txt", header = FALSE, sep = ",")
names(new_data) <- c("name", "sex", "n")
new_data <- new_data %>%
  mutate(prop = `n` / sum(new_data$n)) %>%
  mutate(year = 2018)

complete_data <- full_join(new_data,babynames)
```
```{r, message=FALSE, echo=FALSE, warning=FALSE, fig.align='center'}
#Timeseries for "Eug" and "Eugene"
my_name <- complete_data %>%
  filter(str_detect(name, "Eugene") & sex == 'M') %>%
  group_by(year) %>%
  summarise(prop = sum(prop)) %>%
  mutate("Eug" = FALSE)

initials <- complete_data %>%
  filter(str_detect(name, "^Eug") & sex == 'M') %>%
  group_by(year) %>%
  summarise(prop = sum(prop)) %>%
  mutate("Eug" = TRUE)

my_data <- full_join(my_name, initials)
my_name <- my_name %>%
  select(-c("Eug"))
initials <- initials %>%
  select(-c("Eug"))

ggplot(my_data) +
  geom_jitter(mapping = aes(x = year, y = prop, color = Eug)) + 
  labs(x = "Year", y = "Proportion", color = "First Three Letters Only", title = "Popularity of Eug and Eugene since 1880")
```

I used the string "Eugene" to search for my name, as this is the only spelling of my name I am aware of. My name seems to be the most popular name that starts with letters "Eug" as the popularity of names that start with "Eug" and "Eugene" track very closely to each other. 

```{r, message=FALSE, echo=FALSE, warning=FALSE, fig.align='center'}
#Stats on Eugene
pre_me <- my_name %>%
  filter(year == 1980)
birth_year <- my_name %>%
  filter(year == 1999)
post_me <- my_name %>%
  filter(year == 2018)

pop_data <- tribble(
  ~"Year", ~"Proportion",
  pre_me$year, pre_me$prop,
  birth_year$year, birth_year$prop,
  post_me$year, post_me$prop,
)

ggplot(pop_data) +
  geom_col(mapping = aes(x = Year, y = Proportion, fill = Proportion)) + 
  geom_text(mapping = aes(label = Year, y = Proportion/2, x = Year), color = 'white')
```

My name has become less popular since I was born, and its popularity continues to decline every year. In 1980 the popularity was `r pre_me$prop` in my birth year 1999: `r birth_year$prop` and in 2018: `r post_me$prop`. 


## Emily Archambault
 
```{r, message=FALSE, echo=FALSE, warning=FALSE}
emily <- totalnames %>% 
  filter(str_detect(name, "^E[m]+[ai][l]+[iey]+$")) %>%
  filter(sex == "F")
first3 <- totalnames %>%
  filter(str_detect(name, "^Emi")) %>%
  group_by(year)%>%
  filter(sex == "F")

ggplot() + geom_col(data = first3, mapping = aes(x = year, y = prop)) +
  geom_smooth(data = emily, mapping = aes(x = year, y = prop), se = FALSE, color = "red") + 
  annotate("text", x = 1960, y = 0.0013, label = "Emily", color = "red") +
  ggtitle("Popularity of All Female 'Emi...' Names vs Popularity of 'Emily' ")

emily18 <- emily %>%
  filter(year == "2018")
pop18 <- sum(emily18$prop)

emily01 <- emily %>%
  filter(year == "2001")
pop01 <- sum(emily01$prop)

same3 <- first3 %>%
  filter(year == "2018")
pop3 <- sum(same3$prop)
```
The regular expression I used to match the various spellings of my name is "^E[m]+[ai][l]+[iey]+$".
The popularity of my name in 2018 is `r pop18`, compared to when I was born in 2001 which was `r pop01`, a difference of
0.01107516. The popularity of all (female) names starting with "Emi" in 2018 is `r pop3`, which is 0.001304714 greater than
the proportion of my name alone.


## Jingzhi Zhang

```{r, message=FALSE, echo=FALSE, warning=FALSE}
katherine <- totalnames %>%
  filter(str_detect(name, "Katherine")) %>%
  filter(sex == 'F')
kat <- totalnames %>%
  filter(str_detect(name, "^Kat")) %>%
  group_by(year) %>%
  filter(sex == 'F')
ggplot()+
  geom_col(data = kat, mapping = aes(x = year, y = prop))+
  geom_smooth(data = katherine, mapping = aes(x = year, y = prop), se = FALSE, color = "green")+
  annotate("text", x = 1950, y = 0.0045, label = "Katherine", color = "green")
  labs(x = "Year", y = "Proportion", title = "Popularity of Female Kat and Katherine")
  
katherine2018 <- katherine %>%
  filter(year == 2018)
popular2018 <- sum(katherine2018$prop)

katherine1998 <- katherine %>%
  filter(year == 1998)
popular1998 <- sum(katherine1998$prop)

kat2018 <- kat %>%
  filter(year == 2018)
pop2018 <- sum(kat2018$prop)

kat1998 <- kat %>%
  filter(year == 1998)
pop1998 <- sum(kat1998$prop)

```

I search my name by finding the string "Katherine" directly. I found that the popularity of name "Katherine" in 2018 is `r popular2018` and the popularity of it in 1998 which is the year that I was born is `r popular1998`. From the differences of number, it seems like "Katherine" used to be a popular name but less popular at present. And the popularity of "Kat" in 2018 is `r pop2018`, compared to 1998 which is `r pop1998`, this name is much less popular than it used to be.

## Johnny Tamanaha

```{r, message = FALSE, echo = FALSE, warning = FALSE}
joh <- babynames_all %>%
  filter(str_detect(name, "^Joh")) %>%
  group_by(year) %>%
  summarize(joh_prop = sum(prop))
johnny_full <- babynames_all %>%
 filter(str_detect(name, "Jo[ah]?[n]+[ey]+$")) %>%
  group_by(year) %>%
  summarize(johnny_prop = sum(prop))
all_joh <- full_join(joh, johnny_full, by="year")
all_joh <- all_joh %>%
  gather(key=name_type, value=proportion, 2:3)

ggplot(all_joh) +
  geom_point(mapping=aes(x=year, y=proportion, color=name_type)) +
  labs(title="Proportion of 'Johnny' Variants and Names that Start with 'Joh'", color="Name Type") +
  xlab("Year") +
  ylab("Proportion") +
  scale_color_discrete(labels=c("Start with 'Joh'", "'Johnny' Variants"))
```

The regular expression I used for my name was ^Jo[ah]?[n]+ie[y]*$. There has been a small steady decline of Johnnys since my birth year. Other names that start with Joh have been on a heavy decline from 1880 to 2018, so my name has become relatively more popular in comparison to those names. 

## kaiyuan Fan

```{r, message = FALSE, echo = FALSE, warning = FALSE}
Fra <- totalnames %>%
  filter(str_detect(name, "^Fra"))%>%
  filter(sex =="M") %>%
  group_by(year) 
Frank <- totalnames %>%
  filter(str_detect(name, "^Frank$"))%>%
  filter(sex == "M") %>%
  group_by(year)

ggplot()+
  geom_col(Fra, mapping= aes(year, prop), fill ="#393f87" ) +
  geom_smooth(Frank, mapping = aes(year, prop), color = "#aa68b3", se = FALSE)+
  labs(x = "Year", y = "Proportion", title = "Popularity of Fra and Frank since 1880 to 2018") +
  theme(text = element_text(face = "plain",color = "#FF63B8", size = 13, lineheight =1,  hjust = 0.5), plot.title = element_text (hjust = 0.1,color="#8C57A6", size =15))
  
Frank2018 <- Frank %>%
   filter(year == "2018")
Frank1998 <- Frank %>%
   filter(year == "1998")
Fra2018 <- Fra %>%
   filter(year == "2018") %>%
  summarise(sum(prop))
```

I use "Frank" to be regular expression to find my name in dataset.
The popularity of my name in 1998 is `r Frank1998$prop`. However, the popularity of my name in 2018 is `r Frank2018$prop`. Obviously, the popularity declined over years which shows on the plot above. 
The popularity of all (male) names starting with "Fra" in 2018 is `r Fra2018`, which is 0.001168795. It is greater than
the proportion of my name alone.

### Ariel and Rachel Regexs

```{r, message = FALSE, echo = FALSE, warning = FALSE}
arielrachel <- totalnames %>%
  filter(str_detect(name, "A[r]+[iye][ea][l][le]*$") | str_detect(name, "^R[ae][aey]*[c]h[aey]+[l]+e?$")) %>%
  filter(year %in% list(1973, 1988, 1990, 2018)) %>%
  filter(sex == "F")
  
arielrachelversions <- arielrachel %>%
  group_by(year) %>%
  group_by(year, initial=str_sub(name, 1,1))%>%
  summarise(variants = n())

arielrachelprop <- arielrachel %>%
  group_by(year) %>%
  summarise(prop_A_R = sum(prop))

arielrachelversions
arielrachelprop
```

The first table shows the number of versions of the names Ariel or Rachel for each year, and the second table shows the chances that a girl would have either of those names for each year.

### The Little Mermaid Effect

```{r, message = FALSE, echo = FALSE, warning = FALSE}
female_name_changes <- babynames_all %>%
  filter(sex=="F") %>%
  filter(year==1988 | year==1990) %>%
  spread(key=year, value=prop) %>%
  filter(!is.na(`1988`) | !is.na(`1990`)) %>%
  mutate(`1988`=ifelse(is.na(`1988`), 0, `1988`), `1990`=ifelse(is.na(`1990`), 0, `1990`)) %>%
  mutate(difference = `1990` - `1988`)
ariel_change <- female_name_changes %>%
  filter(str_detect(name, "A[r]+[iey][ea][l][le]*$")) %>%
  group_by(sex) %>%
  summarize(total_diff = sum(difference)) 
ggplot(female_name_changes) +
  geom_histogram(aes(difference), bins=1000) +
  geom_vline(xintercept=ariel_change$total_diff, color="Red") +
  labs(title="Change in Proporiton of all Names from 1988 to 1990") +
  xlab("Difference in Proporiton") +
  ylab("Count") +
  ylim(0, 300) +
  xlim(-0.01, 0.01)
```

### Your Names

```{r, message = FALSE, echo = FALSE, warning = FALSE}
all_prop_2018 <- babynames_all %>%
  filter(str_detect(name, "Jo[ah]?[n]+[ey]+$") | str_detect(name, "^E[m]+[ai][l]+[iey]+$") | str_detect(name, "Eugene") | str_detect(name, "Katherine") | str_detect(name, "Frank")) %>%
  filter(year==2018) %>%
  summarize(total_prop = sum(prop)) %>%
  mutate(year = 2018)
all_percentile_2018 <- babynames_all %>%
  mutate(lower_percentile = prop <= all_prop_2018$total_prop) %>%
  summarize(percentile = sum(lower_percentile)/n()) %>%
  mutate(year = 2018)

all_prop_1999 <- babynames_all %>%
  filter(str_detect(name, "Jo[ah]?[n]+[ey]+$") | str_detect(name, "^E[m]+[ai][l]+[iey]+$") | str_detect(name, "Eugene") | str_detect(name, "Katherine") | str_detect(name, "Frank")) %>%
  filter(year==1999) %>%
  summarize(total_prop = sum(prop)) %>%
  mutate(year = 1999)
all_percentile_1999 <- babynames_all %>%
  mutate(lower_percentile = prop <= all_prop_1999$total_prop) %>%
  summarize(percentile = sum(lower_percentile)/n()) %>%
  mutate(year = 1999)

all_prop_1980 <- babynames_all %>%
  filter(str_detect(name, "Jo[ah]?[n]+[ey]+$") | str_detect(name, "^E[m]+[ai][l]+[iey]+$") | str_detect(name, "Eugene") | str_detect(name, "Katherine") | str_detect(name, "Frank")) %>%
  filter(year==1980) %>%
  summarize(total_prop = sum(prop)) %>%
  mutate(year = 1980)
all_percentile_1980 <- babynames_all %>%
  mutate(lower_percentile = prop <= all_prop_1980$total_prop) %>%
  summarize(percentile = sum(lower_percentile)/n()) %>%
  mutate(year = 1980)

prop_compare_1999_2018 <- rbind(all_prop_1999, all_prop_2018)
ggplot(prop_compare_1999_2018) +
  geom_col(aes(year, total_prop)) +
  labs(title="Group Names Proportion in 1999 and 2018") +
  xlab("Year") +
  ylab("Proportion")

prop_compare_1999_1980 <- rbind(all_prop_1999, all_prop_1980)
ggplot(prop_compare_1999_1980) +
  geom_col(aes(year, total_prop)) +
  labs(title="Group Names Proportion in 1980 and 1999") +
  xlab("Year") +
  ylab("Proportion")

percentile_compare_1999_2018 <- rbind(all_percentile_1999, all_percentile_2018)
ggplot(percentile_compare_1999_2018) +
  geom_col(aes(year, percentile)) +
  labs(title="Group Names Percentile in 1999 and 2018") +
  xlab("Year") +
  ylab("Percentile")

percentile_compare_1999_1980 <- rbind(all_percentile_1999, all_percentile_1980)
ggplot(percentile_compare_1999_1980) +
  geom_col(aes(year, percentile)) +
  labs(title="Group Names Percentile in 1980 and 1999") +
  xlab("Year") +
  ylab("Percentile")

```

```{r, echo = FALSE}
female_name_changes <- babynames_all %>%
  filter(sex=="F") %>%
  filter(year==1988 | year==1990) %>%
  spread(key=year, value=prop) %>%
  filter(!is.na(`1988`) & !is.na(`1990`)) %>%
  mutate(`1988`=ifelse(is.na(`1988`), 0, `1988`), `1990`=ifelse(is.na(`1990`), 0, `1990`)) %>%
  mutate(difference = `1990` - `1988`)
ariel_change <- female_name_changes %>%
  filter(str_detect(name, "A[r]+[iey][ea][l][le]*$")) %>%
  group_by(sex) %>%
  summarize(total_diff = sum(difference)) 
ggplot(female_name_changes) +
  geom_histogram(aes(difference), bins=100) 
  geom_vline(xintercept=ariel_change$total_diff)
```

### Individual Summaries

# Emily Archambault
I created an individual plot showing a comparison between the popularity of my name and all other names starting with "Emi". I also did the Ariel/Rachel regex section.

# Johnny Tamanaha
Johnny Tamanaha: Along with my individual section I did the plot for the changes in female names from 1988 to 1990. I also found the percentiles of all our names in 1980, 1999 and 2018.

# Kaiyuan Fan
I create a plot over time (1880-2018) of the proportion of names in a given year that start with the first three letters of your name. And I find the populariy of my name declined over years.

# Eugene Miller

# jingzhi Zhang
I create a plot of proportion of names with "katherine"(female) and start with the first three letters of your name over years. 